---
title: "Example Final Paper"
author: "Nathaniel Phillips"
date: "5/9/2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

# Note

The following is a template you can use for writing your final paper. In your paper, you should include all of your raw code in your document.


# Dataset Description

I obtained the data from the international chicken data repository. I don't know how the data were originally collected. The main columns in the data were as follows: *weight*, the weight of chickens in grams, *Time* the age in weeks of the chick at the time of measurement, *Chick* a unique number for each chicken, and *Diet* the diet given to the chick.

# Questions

I will answer the following questions in my paper.

1. How did the chicken weights generally change over time?
2. Was there a difference in the the average chicken weights as a result of the different diets?
3. ...
4. ...
5. ...

# Analyses

```{r, echo = TRUE}
# Task 1: Load data
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 1: Load data

ChickWeight <- read.table("data/ChickWeight.txt", 
                          sep = "\t", 
                          header = TRUE)
```

Summary statistics from the data are presented in the following table.

```{r, echo = TRUE}
# Task 2: Summary statistics from a dataframe
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 2: Summary statistics from a dataframe
summary(ChickWeight)
```

The data had 4 columns: weight, Time, Chick, and Diet:

```{r, echo = TRUE}
# Task 3: Printing variable names
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 3: Printing variable names

names(ChickWeight)
```

The data for Diet were originally coded as numbers, I recoded the Diet data as string variables

```{r, echo = TRUE}
# Task 4: Recoding a variable
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 4: Recoding a variable

ChickWeight$Diet.s <- NA

ChickWeight$Diet.s[ChickWeight$Diet == 1] <- "bacon"
ChickWeight$Diet.s[ChickWeight$Diet == 2] <- "cheese"
ChickWeight$Diet.s[ChickWeight$Diet == 3] <- "tofu"
```


```{r, echo = TRUE}
# Task 5: Calculate simple summary statistics
# INSERT CODE HERE
```

The mean weight of chickens across all data was `r round(mean(ChickWeight$weight), 2)`, the median weight was `r round(median(ChickWeight$weight), 2)` and the standard deviation was `r round(sd(ChickWeight$weight), 2)`.

A table of frequencies showing how many observations there were for each diet is displayed in the following table:

```{r, echo = TRUE}
# Task 6: Print a table
# INSERT CODE HERE
```

```{r, echo = FALSE, results = 'asis'}
# Task 6: Print a table

chick.table <- data.frame(table(ChickWeight$Diet))
names(chick.table) <- c("Diet", "Frequency")
papaja::apa_table(chick.table, caption = "Number of chicks on each diet")
```


```{r, echo = TRUE}
# Task 7: Count outliers
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 7: Count outliers

# Logical vector of outliers
outlier.log <- with(ChickWeight, 
                      (weight > mean(weight) + 3 * sd(weight)) | 
                      (weight < mean(weight) - 3 * sd(weight))) 

# How many are there?
outlier.sum <- sum(outlier.log)
```

To see if there were any outliers in the weight data, I counted how many chicks had weights greater than 3 standard deviations above the mean, or less than 3 standard deviations below the mean. Using this procedure, I counted `r outlier.sum` outliers.

A scatterplot showing the relationship between time and weight is shown in the following figure


```{r, echo = TRUE}
# Task 8: Scatterplot with regression line.
# INSERT CODE HERE
```

```{r ,fig.cap="Scatterplot of chicken weights over time.", echo = FALSE}
# Task 8: Scatterplot with regression line.

plot(x = ChickWeight$Time,
     y = ChickWeight$weight,
     xlab = "Time (in weeks)",
     ylab = "Weight (in grams)")

grid()

mod <- lm(weight ~ Time, 
          data = ChickWeight)

abline(mod, col = "blue", lty = 2)
```


A histogram of the weight data are presented in the next figure

```{r, echo = TRUE}
# Task 9: Histogram
# INSERT CODE HERE


# Task 20: Custom Function: my.hist()
# Insert code here
```

```{r, echo = FALSE, fig.cap = "Distribution of weights across all data points."}
# Task 9: Histogram
hist(ChickWeight$weight, 
     main = paste0("Chicken Weights (mean = ", round(mean(ChickWeight$weight), 2), ")"), 
     xlab = "Weight (in grams)",
     col = "papayawhip")

abline(v = mean(ChickWeight$weight))
```

Histograms separately for each diet are displayed in the next figure


```{r, echo = TRUE}
# Task 20: Loop
# INSERT CODE HERE
```

```{r, echo = FALSE, fig.cap = "Histograms of the distribution of weights across time for each diet. Vertical lines are means."}
par(mfrow = c(2, 2))

for(Diet.i in 1:4) {
  
  hist(ChickWeight$weight[ChickWeight$Diet == Diet.i], 
       main = paste("Diet =", Diet.i), xlim = c(0, 400), xlab = "weight")
  
  abline(v = mean(ChickWeight$weight[ChickWeight$Diet == Diet.i]), lty = 2)
  
}
```



A pirateplot showing the relationship between diet and weight is shown here:

```{r, echo = TRUE}
# Task 10: pirateplot
# INSERT CODE HERE
```

```{r, echo = FALSE, fig.cap= "Pirateplot showing the distribution of chicken weights by diet. Horizontal lines show means while white boxes show Bayesian 95\\% highest density intervals."}
# Task 10: pirateplot

yarrr::pirateplot(weight ~ Diet,
                  data = ChickWeight)
```

The mean weight of chicks on each diet is shown in the following table:


```{r, echo = TRUE}
# Task 11: Descriptive statistics across groups
# INSERT CODE HERE
```

```{r, echo = FALSE, results = 'asis'}
# Task 11: Descriptive statistics across groups

diet.agg <- aggregate(weight ~ Diet, 
                      data = ChickWeight, 
                      FUN = mean)

names(diet.agg) <- c("Diet", "Mean Weight")

papaja::apa_table(diet.agg, 
                  caption = "Mean weights of chickens on each diet")
```



```{r, echo = TRUE}
# Task 12: 1 sample t-test
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 12: 1 sample t-test

weight.ttest <- t.test(ChickWeight$weight, 
                       mu = 100)

# Print the result
weight.ttest
```

A one sample t-test comparing the weights of chickens to a null hypothesis of 100 was significant `r papaja::apa_print(weight.ttest)$full_result`. The mean weight of chickens was significantly larger than 100 grams.


```{r, echo = TRUE}
# Task 13: t-test with subset
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 13: t-test with subset

weight.subset.ttest <- t.test(weight ~ Diet,
                              data = ChickWeight,
                              subset = Diet %in% c(1, 2))

# Print the result
weight.subset.ttest
```

A two sample t-test comparing the weights of chickens between diets 1 and 2 was significant `r papaja::apa_print(weight.subset.ttest)$full_result`, the weights of chickens was significantly higher in diet 2 compared to diet 1.


```{r, echo = TRUE}
# Task 14: correlation test
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 14: correlation test

weight.time.cortest <- cor.test(~ Time + weight,
                              data = ChickWeight)

# Print the result
weight.time.cortest
```


A correlation test detecting a relationship between time and weight was significant `r papaja::apa_print(weight.time.cortest)$full_result`, as time increased, the weight of chickens increased.

```{r, echo = TRUE}
# Task 15: correlation test with subset
# INSERT CODE HERE
```


```{r, echo = FALSE}
# Task 15: correlation test with subset

weight.time.diet2.cortest <- cor.test(~ Time + weight,
                              data = ChickWeight,
                              subset = Diet == 2)

# Print the result
weight.time.diet2.cortest
```

A correlation test detecting a relationship between time and weight only for chickens on diet 2 was significant `r papaja::apa_print(weight.time.diet2.cortest)$full_result`, as time increased, the weight of chickens on diet 2 increased.

```{r, echo = TRUE}
# Task 16: Chi-Square test
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 16: Chi-Square test

diet.csqtest <- chisq.test(table(ChickWeight$Diet))

# Print the result
diet.csqtest
```

To see if there was a significant difference in the number of chickes on each diet. I performed a chi-square test. The test was significant `r papaja::apa_print(diet.csqtest, n = nrow(ChickWeight))$statistic`, indicating that chickens were not equally distributed amongst the diets.



```{r, echo = TRUE}
# Task 17: ONE-WAY ANOVA
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 17: ONE-WAY ANOVA

diet.anova <- aov(weight ~ factor(Diet), 
                  data = ChickWeight)

# Print the result
diet.anova
summary(diet.anova)

# Post hoc tests
diet.tukey <- TukeyHSD(diet.anova)
```

To see if there was a significant difference in weights between diets, I performed a one-way ANOVA. The test was significant `r papaja::apa_print(diet.anova)$full_result$Diet`, indicating that there was a significant difference between diets. Post-hoc tests showed significant differences between Diets 1 and 3 (diff = 40.30, p < .01), and Diets 1 and 4 (diff = 32.62, p < .01).



```{r, echo = TRUE}
# Task 18: TWO-WAY ANOVA
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 18: TWO-WAY ANOVA

diet.2anova <- aov(weight ~ factor(Diet) + factor(Time), 
                  data = ChickWeight)

# Print the result
diet.2anova
summary(diet.2anova)

# Post hoc tests
diet.2tukey <- TukeyHSD(diet.2anova)
```

To see if there was a significant difference in weights between diets and time points, I performed a two-way ANOVA. The effect of both diet `r papaja::apa_print(diet.2anova)$full_result$factorDiet` and time `r papaja::apa_print(diet.2anova)$full_result$factorTime` were significant. Post-hoc tests showed significant differences between all Diets except for 4 and 3 (diff = -7.69, p = 0.34). There were significant differences between almost all pairs of time periods. 


```{r, echo = TRUE}
# Task 19: REGRESSION
# INSERT CODE HERE
```

```{r, echo = FALSE}
# Task 19: REGRESSION

weight.lm <- lm(weight ~ Time, 
                  data = ChickWeight)

# Print the result
weight.lm
summary(weight.lm)
```

To see if time was related to weight, I regressed weight on time. Results showed a significant positive effect of time `r papaja::apa_print(weight.lm)$full_result$Time`, `r papaja::apa_print(weight.lm)$full_result$modelfit$r2`


# Conclusion

The two most important results were that chickens gain weight over time, and Diet 3 lead to the highest weights while Diet 1 lead to the lowest weights.


