---
title: "Random data for final project instructions"
author: "Nathaniel"
date: "11/27/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


To get random data for your final project, you will


### Step 1: Look at the GenerateFinalProject.R file

Open a web browser and go to [https://raw.githubusercontent.com/ndphillips/IntroductionR_Course/master/assignments/finalproject/GenerateFinalProject.R](https://raw.githubusercontent.com/ndphillips/IntroductionR_Course/master/assignments/finalproject/GenerateFinalProject.R). As you'll see, this page contains R code which defines a function called `final.data.fun()`. Look through the code and try to understand generally how it works.

### Step 2: Get the GenerateFinalProject() function

In R Studio, create a new R script. Then, run the following code. This code will evaluate all of the code in the `generate_finaldata.R` file in your current R session. After you do this, you should now have the `GenerateFinalProject()` function in your current R session.

```{r}
# Evaluate all of the generate_finaldata.R code in your current R session.

source(file = url("https://raw.githubusercontent.com/ndphillips/IntroductionR_Course/master/assignments/finalproject/GenerateFinalProject.R"))
```

### Step 3: Use GenerateFinalData() to create your data

when you evaluate the `GenerateFinalProject()` function, it will generate the data, and save the result as a file called `priming_XXX.txt`, where XXX is your three digit *id*.

The main argument to `GenerateFinalProject()` is `my.id`, which is a three digit id. When you specify a unique value of `my.id`, the dataset the function generates will always be the same.

To make sure that every student has a different dataset, you should set the `my.id` argument to the last three digits of your university matriculation number (If you don't have this number, ask me in class and I can look it up for you). If you don't have a matriculation number, you can use the last three digits of your mobile phone number.


For example, if your id was 123, you could run the following

```{r}

GenerateFinalProject(my.id = 123)

```

